\vspace{1.0cm} \LARGE \bf Supporting Text: \\ Revealing the competitive dynamics of influenza viruses through evolutionary cartography

### 2013-01-24 flux_si.tex
Generalize diffusion coefficient derivation per Marc's suggestion.

The distance $d$ traveled by two-dimensional diffusion is not a linear function of time interval $t$, and thus the `rate' of diffusion cannot be calculated following $r=d/t$.
Assume we have a one-dimensional diffusion starting from $x_0 = 0$, so that
\begin{equation}
	X_t \sim \normal(0, \sigma^2 t),
\end{equation}
where $\sigma$ represents the diffusion volatility, and in this case, the expected squared displacement follows
\begin{equation}
	\mathrm{E}[X_t^2] = \sigma^2 t.
\end{equation}
In moving to a two-dimensional diffusion
\begin{equation}
	(X_t, Y_t) \sim \normal \left( \twomatrix{0}{0}, \fourmatrix{\sigma^2 t}{0}{0}{\sigma^2 t} \right),
\end{equation}
a factor of 2 is included when calculating expected squared displacement
\begin{equation}
	\mathrm{E}[X_t^2 + Y_t^2] = 2 \sigma^2 t.
\end{equation}
Such diffusions are often parameterized with diffusion coefficient $D$ rather than volatility parameter $\sigma$.
In this case $D=\sigma^2/2$ and 
\begin{equation}
	\mathrm{E}[X_t^2 + Y_t^2] = 4 D t.
\end{equation}

>>>>

The distance $d$ traveled by a diffusion is not a linear function of time interval $t$, and thus the `rate' of diffusion cannot be calculated following $r=d/t$.
Assume we have a one-dimensional diffusion starting from $x_0 = 0$, so that
\begin{equation}
	X_t \sim \normal(0, \sigma^2 t),
\end{equation}
where $\sigma$ represents the diffusion volatility, and in this case, the expected squared displacement follows
\begin{equation}
	\mathrm{E}[X_t^2] = \sigma^2 t.
\end{equation}
In moving to a two-dimensional diffusion
\begin{equation}
	(X_t, Y_t) \sim \normal \left( \twomatrix{0}{0}, \fourmatrix{\sigma_x^2 t}{\rho \sigma_x \sigma_y t}{\rho \sigma_x \sigma_y t}{\sigma_y^2 t} \right),
\end{equation}
where $\sigma_x$ represents volatility in dimension 1, $\sigma_y$ represents volatility in dimension 2 and $\rho$ represents correlation,   expected squared displacement is equal to the sum of squared displacements in each dimension
\begin{equation}
	\mathrm{E}[X_t^2 + Y_t^2] = \sigma_x^2 t + \sigma_y^2 t.
\end{equation}
Such diffusions are often characterized with diffusion coefficient $D$.
In this case
\begin{equation}
	D = \frac{\sigma_x^2 + \sigma_y^2}{4},
\end{equation}
so that expected squared displacement follows from $D$ alone
\begin{equation}
	\mathrm{E}[X_t^2 + Y_t^2] = 4 D t.
\end{equation}
